---
title: "Visualization and EDA"
subtitle: "Part II"
output: github_document
---

# Setup

```{r, setup}
library(tidyverse)
library(patchwork)
library(hexbin)
```

# Data Input
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Remember this plot?
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)
```

## Add Labels

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  )
```

## Scales
Defining the x and y scales, with transformation.

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "log"
  )
```

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 300)
    )
```

Viridus package
```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```

Themes

Shift the legend position

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

Change the overall theme.

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal()
```

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature Plot",
    x = "Minimum Daily Temperature (C)",
    y = "Maximum Daily Temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  ggthemes::theme_fivethirtyeight()
```

## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6
  fig.width = .6,
  out.width = "90%"
)

theme_set(
  ggthemes::theme_fivethirtyeight() + theme(legend.position = "bottom")
  )

options(
  ggplot2.continuous.colour = "viridis"
  ggplot2.continuous.colour = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in geom

Here you can see how one data in the geom_point is DIFFERENT from the geom_line by creating a second/ separate dataset.  
```{r}
central_park =
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki =
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
    geom_point() +
    geom_line(data = central_park)
```

## `patchwork`

Remember faceting?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

What happens when you want multi panel plots, but can't facet?

```{r}
tmax_tmin_p =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

# Data manipulation

With releving.  In order to change the categoricals on the x,you should use data manipulation first by reodering.  
Control your factors!

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Waikiki_HA"))
  ) %>% 
  ggplot(aes(y = tmax, x = name, fill = name)) +
  geom_violin(alpha = .5)
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5)
```

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

