---
title: "Visualization and EDA"
subtitle: "Part I"
output: github_document
---

# Setup

```{r, setup}
library(tidyverse)
library(ggridges)
library(hexbin)
```

# Data Input
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots!!

Create a first scatterplot ever.

```{r}
ggplot(weather_df, aes(x = tmin, y  = tmax)) + 
  geom_point()
```

New approach, same plot

```{r}
weather_df  %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Save and edit a plot object.

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot...  
Start with the same one and make it fancy
```{r}
weather_df  %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

What about the aes placement?
So if you move the aes to the second line, it only defines the colors to the point, but not the smooth
```{r}
weather_df  %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth()
```

Let's facet!
```{r}
weather_df  %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se= FALSE) +
  facet_grid(. ~ name)
```

Coloring, customization

```{r}
weather_df  %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .2, size = .3) +
  geom_smooth(se= FALSE, size = 2) +
  facet_grid(. ~ name)
```

```{r}
weather_df  %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) + 
  geom_point() +
  geom_smooth(se= FALSE, size = 2) +
  facet_grid(. ~ name)
```

Let's combine elements and try a new plot.

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

How many goems have to exist?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = 0.3)
```

Univariate plots

Histograms are really great.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

A bit confusing, let's facet it?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Let's try a new geometry.
Density is like a histogram, filled in with smooth edges.
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5)
```

What about box plots?
Note that there is a categorical variable on the x axis!  This works for boxplots.  

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Trendy plots :-)

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin()
```

Add a statistic
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "median")
```

Trendy Ridgeplots :-)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name, fill = name)) + 
  geom_density_ridges()
```

Let's save a scatterplot.

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)

ggsave("weather_plot.pdf", weather_plot)
```

What about embedding?
```{r, fig.width=6, fig.height=3}
weather_plot 
```

```{r, fig.width=12, fig.height=6}
weather_plot 
```


