---
title: "Data Manipulation"
output: github_document
---

# Setup

```{r, setup}
library(tidyverse)
```

## Load in FAS_Litters dataframe

```{r}
litters_df = read_csv(file = "./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## select

Choose some columns and not others.

```{r}
select(litters_df, group, litter_number, gd0_weight:gd_of_birth)
```

If we want to remove a variable.
```{r}
select(litters_df, -group)
```

Rename a column and select it
```{r}
select(litters_df, GROUP = group)
```

Just rename a column.

```{r}
rename(litters_df, GROUP = group)
```

Select helpers

```{r}
select(litters_df, starts_with("gd"))
```

Relocate to the beginning and not do anything else
```{r}
relocate(litters_df, litter_number)
```
## Filter

Filter operates on ROWS 

```{r}
filter(litters_df, gd0_weight <22)
```

Make it equal to 22
```{r}
filter(litters_df, gd0_weight ==22)
```

Make it everything except when it equals 22

```{r}
filter(litters_df, !(gd0_weight == 22))
```

Filter and select multiple character rows 

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## Mutate

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight
  )
```

## Arrange

```{r}
arrange(litters_df, pups_born_alive)
```
## %>%

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```

