---
title: "Exploratory Analysis"
subtitle: "Part III"
output: github_document
---

# Setup

```{r, setup}
library(tidyverse)
library(patchwork)
library(hexbin)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6,
  out.width = "90%"
)

theme_set(
  ggthemes::theme_fivethirtyeight() + theme(legend.position = "bottom")
  )

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.colour = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## group_by

Here now we have defined three groups, a new grouping structure.

```{r}
weather_df %>% 
  group_by(name, month)
```

And now ungroup at the end.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup()
```

## counting things

Counting month observations
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())
```

Group by name
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

We can use count

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```

Other helpful counters

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

## A digression 2x2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE ~ "")) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

Another way to do it!
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE ~ "")) %>% 
  janitor::tabyl(name, cold)
```

## General Summaries
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    media_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe!  
The `group by` and `summarize` are creating a nice new dataset for us to plot.  

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    media_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Suppose you want to summarize many columns in the same way.
Use the across function!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

Just a reminder, sometimes your data results are easier to read in another format!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

## group_by and mutate

So here we have grouped, and then created a new column with the mean and now plotted the values as a difference from the mean
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y= centered_tmax, color = name)) +
  geom_point()
```

What about window functions?
So here we have grouped by name and then assigned relative ranks among the tmax variable, then we looked at the tmax of every month.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    temp_rank = min_rank(tmax)) %>% 
  filter(temp_rank ==1)
```

lag
Here we are taking the values, and then it creates a new column with the value, down a row, or however many you want. 
Does yesterday's temp predict today's?

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = lag(tmax))
```

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(temp_change = tmax - lag(tmax)) %>% 
  summarize(
    temp_change_max = max(temp_change, na.rm = TRUE),
    temp_change_sd = sd(temp_change, na.rm = TRUE))
```
